---
title: "Comparing Current well-being with its demographic dataflows"
---

```{r}
#| label: setup
#| include: false

source(here::here("R", "COMPARE_DATAFLOWS.R"))

comparisons <- list(
  Age = readRDS(
    here::here(
      "data",
      "processed",
      "comparison_age.rds"
    )
  ),
  Sex = readRDS(
    here::here(
      "data",
      "processed",
      "comparison_sex.rds"
    )
  ),
  `Education level` = readRDS(
    here::here(
      "data",
      "processed",
      "comparison_education.rds"
    )
  )
)

overview <- dplyr::bind_rows(
  lapply(
    names(comparisons),
    function(comparison_name) {
      COMPARISON_OVERVIEW(
        comparisons[[comparison_name]]
      ) |>
        dplyr::mutate(
          Comparison = comparison_name,
          .before = 1
        )
    }
  )
)
```

## Question

The OECD disseminates *Current well-being* alongside specialized views by age, sex, and education level [@oecd_data_explorer_hows_life_2026]. Do these demographic dataflows contain distinct non-missing `OBS_VALUE` observations, or do their values already appear in the main *Current well-being* dataflow when records are matched on their shared SDMX keys?

The comparison is deliberately narrow: it evaluates record-key overlap and `OBS_VALUE` agreement. It does not establish equivalence of observation-status flags, other SDMX attributes, metadata, or presentation conventions.

## Comparison Design

The three specialized dataflows use the same core demographic dimensions as *Current well-being*, but constrain the demographic codes differently. In the current SDMX structures, the specialized scopes are:

| Comparison | Age | Sex | Education level |
|---|---|---|---|
| **Current well-being by age** | **`MID`, `OLD`, `YOUNG`** | `_T` (total) | `_T` (total) |
| **Current well-being by sex** | `_T` (total) | **`F`, `M`** | `_T` (total) |
| **Current well-being by education** | `_T` (total) | `_T` (total) | **`ISCED11_1`, `ISCED11_2_3`, `ISCED11_5T8`** |

: Demographic codes represented by the three specialized dataflows in the current SDMX structures. {#tbl-demographic-scope}

The project applies the same comparison procedure to all three specialized views. Records are matched on the canonical SDMX dimensions shared by the parent and specialized datasets: `REF_AREA`, `MEASURE`, `UNIT_MEASURE`, `AGE`, `SEX`, `EDUCATION_LEV`, `DOMAIN`, and `TIME_PERIOD`. A full join preserves matched keys as well as keys with a non-missing value on only one side. `OBS_VALUE` is compared using an absolute tolerance of `1e-10`.

The resulting statuses have the following meanings:

- **Exact match**: both sides contain non-missing values that agree within tolerance
- **Different value**: both sides contain non-missing values that differ by more than the tolerance
- **Parent only**: only *Current well-being* contains a non-missing value for the joined key
- **Child only**: only the specialized dataflow contains a non-missing value for the joined key
- **Both missing**: neither joined value is available

Importantly, **parent-only and child-only are value-availability classifications, not independent tests of row presence**. A record that is present with a missing `OBS_VALUE` is therefore treated the same as a missing joined value for this classification.

::: {.data-note}
### [How to interpret parent-only records]{.smallcaps}

The current comparison joins the complete parent snapshot to each specialized snapshot on their shared canonical dimensions. Because the specialized dataflows constrain demographic codes more narrowly than the parent, **parent-only records include parent observations outside the specialized demographic scope** as well as any other one-sided non-missing values.

Accordingly, the exact-match shares below describe the full joined record-key universe. They should not be interpreted as coverage rates for the specialized dataflows' intended demographic scopes.
:::

::: {.context-note}
### [Subgroup-data caveats]{.smallcaps}

OECD notes that indicators in the specialized age and education views are **banded when the sample size is below 1,000 observations** [@oecd_cwb_age_2026; @oecd_cwb_edu_2026]. The Data Monitor also notes that some subgroup estimates pool multiple years, so a displayed period should not automatically be interpreted as a single-year estimate [@oecd_wellbeing_monitor_2026].

These source-level conventions are separate from the record-matching exercise below.
:::

## Results

@tbl-demographic-comparison-summary places the three comparisons side by side. This is the main advantage of treating the exercise as one analysis rather than three separate pages.

```{r}
#| label: tbl-demographic-comparison-summary
#| tbl-cap: "Record-level comparison of Current well-being with its specialized demographic dataflows."

overview |>
  dplyr::mutate(
    `Exact-match share` = sprintf(
      "%.1f%%",
      100 * exact_match / total_records
    )
  ) |>
  dplyr::transmute(
    Comparison,
    `Record keys` = format(
      total_records,
      big.mark = ",",
      scientific = FALSE,
      trim = TRUE
    ),
    `Exact matches` = format(
      exact_match,
      big.mark = ",",
      scientific = FALSE,
      trim = TRUE
    ),
    `Exact-match share`,
    `Parent only` = format(
      parent_only,
      big.mark = ",",
      scientific = FALSE,
      trim = TRUE
    ),
    `Child only` = format(
      child_only,
      big.mark = ",",
      scientific = FALSE,
      trim = TRUE
    ),
    `Different value` = format(
      different_value,
      big.mark = ",",
      scientific = FALSE,
      trim = TRUE
    ),
    `Both missing` = format(
      both_missing,
      big.mark = ",",
      scientific = FALSE,
      trim = TRUE
    )
  ) |>
  knitr::kable()
```

In the frozen project snapshot, all three comparisons report **zero child-only observations, zero value disagreements, and zero both-missing joined keys**. Every non-missing value available through the specialized demographic dataflows therefore has a corresponding non-missing value in *Current well-being* and agrees with it within the specified tolerance.

The reverse does not hold: *Current well-being* contains many parent-only values in each full join. Because the parent includes demographic combinations outside each specialized scope, those counts demonstrate that the parent is broader; they should not be read as missing-data rates for the specialized views.

## Parent-Only Records

The tabs below retain a small record-level audit sample from each comparison. They are useful for seeing what the one-sided parent classifications look like without repeating the same diagnostic structure three times.

::: {.panel-tabset}

### Age

```{r}
#| label: tbl-age-parent-only-records
#| tbl-cap: "First up to 20 records classified as available only in Current well-being in the age comparison."

parent_only_age <- COMPARISON_RECORDS(
  comparison = comparisons$Age,
  status = "Parent only"
)

if (nrow(parent_only_age) == 0L) {
  cat("No parent-only records were detected.")
} else {
  knitr::kable(parent_only_age)
}
```

### Sex

```{r}
#| label: tbl-sex-parent-only-records
#| tbl-cap: "First up to 20 records classified as available only in Current well-being in the sex comparison."

parent_only_sex <- COMPARISON_RECORDS(
  comparison = comparisons$Sex,
  status = "Parent only"
)

if (nrow(parent_only_sex) == 0L) {
  cat("No parent-only records were detected.")
} else {
  knitr::kable(parent_only_sex)
}
```

### Education level

```{r}
#| label: tbl-education-parent-only-records
#| tbl-cap: "First up to 20 records classified as available only in Current well-being in the education-level comparison."

parent_only_education <- COMPARISON_RECORDS(
  comparison = comparisons[["Education level"]],
  status = "Parent only"
)

if (nrow(parent_only_education) == 0L) {
  cat("No parent-only records were detected.")
} else {
  knitr::kable(parent_only_education)
}
```

:::

## Specialized-Only Records and Value Disagreements

No specialized-only non-missing values are detected in the frozen age, sex, or education comparisons. Likewise, no matched keys report `OBS_VALUE` disagreements beyond the `1e-10` tolerance.

This is the central result of the audit: **with respect to the shared canonical keys and non-missing `OBS_VALUE`, the three specialized demographic views add no unique values in the frozen snapshot**. This statement is intentionally narrower than saying that the specialized dataflows are identical to, or literal subsets of, *Current well-being*.

## Interpretation and Limitations

The comparison supports a practical conclusion for this project: the specialized demographic dataflows do not contribute additional non-missing `OBS_VALUE` values beyond those already represented in *Current well-being* on matching keys. Their value lies instead in providing dedicated dissemination views with narrower demographic scopes and their own metadata and reliability conventions.

The result should not be generalized beyond what is actually tested. In particular, the audit does not establish equivalence of observation-status flags, methodological annotations, other attributes, or metadata. It also does not distinguish a genuinely absent row from a present row whose `OBS_VALUE` is missing when assigning the parent-only and child-only categories.

<script src="../assets/site-refinements.js" defer></script>

{{< oecd-rules >}}
